\chapter{Some pictures of spectral sequences}

Some spectral sequences are too hard to compute with, but now that readers know enough about spectral sequences to interpret diagrams, some completely unproven but important pictures might be appropriate to include in the tail of the book.

\section{$E_2$ for the mod $2$ Adams spectral sequence}

\section{$E_2$ for the $MU$-Adams-Novikov spectral sequence}

\section{Slice spectral sequence and the Kervaire invariant}

\section{The chromatic spectral sequence, stabilizer spectral sequences}

The chromatic spectral sequence, using Wilson's BP sampler.  Existence of the Morava stabilizer spectral sequence, maybe pictures of the stabilizer spectral sequence for the $K(1)$- and $K(2)$-local spheres (at $p \ge 5$?).

\section{The EHP spectral sequence}

Michael Donovan already has a start of a flipbook at \url{http://math.mit.edu/~mdono/\_EHPSS.pdf}, which maybe can be cannibalized for this project.  I pray that he hasn't automated this process and beat me to writing this book!

In fact, much of Haynes Miller's notes on the vector fields on spheres problem deals with and can be restated in terms of the EHP spectral sequence.  Maybe this should be upgraded to its own section, so we can see some unstable phenomena?

\section{Devinatz-Hopkins-Smith and the $X(n)$-Adams spectral sequence}

Is this do-able?  All they show are vanishing lines, but it seems like you should be able to draw a little bit...

\section{$\A(2)$}

Borrow Henriques-Hill?
